---
title: "Halloween Candy Practice"
author: "Ava"
date: "2025-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse) 
library(dplyr) 
library(ggplot2) 
library(knitr) 
library(psych)
library(tidycensus)
```

```{r}
candy <- read_csv('candydata.csv', col_names = TRUE)

```

**Select**



```{r}
#Select chocolate, caramel and sugarpercent 
candy %>% select(chocolate,caramel,sugarpercent) %>% 
  arrange(desc(sugarpercent))
```

**Arrange**

```{r}
  candy %>% select(chocolate,caramel,sugarpercent) %>% 
              arrange((sugarpercent)) 

## default is ascedning 
```



**Filter**

How many chocolate?

  candy %>% select(chocolate,caramel,sugarpercent) %>% 
  filter(chocolate == 1)

How many are chocolate and caramel?
```{r}
  candy %>% select(chocolate,caramel,sugarpercent) %>% 
  filter(chocolate == 1, caramel == 1)
```


How many are chocolate or caramel?
```{r}
  candy %>% select(chocolate,caramel,sugarpercent) %>% 
  filter(chocolate == 1 | caramel == 1)
```


**Combine commands - connect with %>% **

Is candy that is pluribus more expensive?

  candy %>% 
      group_by(pluribus) %>% 
      summarise(count=n(),mean = mean(pricepercent))   

```{r}
candy %>% 
  group_by(pluribus) %>% 
  summarise(count=n(),mean = mean(pricepercent))
```

### Practice

Is caramel more expensive?
Yes!
```{r}
candy %>% 
  group_by(caramel) %>% 
  summarise(count=n(),mean = mean(pricepercent))
```

Is there more chocolate or non-chocolate?
Non-Chocolate

```{r}
candy %>% 
  group_by(chocolate) %>% 
  summarise(count=n())
```


What is the winning-est fruit candy?
44.11% 

```{r}
candy %>% 
  group_by(fruity) %>% 
  summarise(count=n(),mean = mean(winpercent))
```


## Plotting

**Basic plots**
  plot(candy$sugarpercent,candy$winpercent)
  

**Fancier plots**

  ggplot(
    data = candy,
    mapping = aes(x = sugarpercent, y = winpercent, color = chocolate)
  ) +
    geom_point()
  
